First, I would like to thank Graham Coop for encouraging me to trust my ideas and find my own path as a scientist. May I grow into being as kind and thoughtful of a mentor and teacher. 

To Coopons past and present, it’s been a great joy and honor to be your peer: Kristin Lee, Doc Edge, Sivan Yair, Matt Osmond, Emily Josephs, Nancy Chen, Vince Buffalo, Katie Ferris, Jeff Groh, Pavitra Muralidhar, Carl Veller, and Jeremy Berg.

I am grateful to many colleagues and friends for supporting my growth as a scientist. Thank you to Jeff Ross-Ibarra for modeling open science and brightening my time at UC Davis with your enthusiasm for Zea evolution, population genetics, and a-maize-ing puns. Thank you to Michael Turelli for treating me like a junior colleague from day one and sharing your love for good science and good wine. Thank you to Annie Schmitt and Chuck Langley for encouraging me to think broadly. To all of the staff in Evolution and Ecology, and especially Sherri Mann, thank you for your kindness and taking care of all the other things so I could focus on my research. 

To all of the CPB students who met my fledgling scientific ideas with perhaps more enthusiasm then they deserved, you are the spark. A special shout-out to my PhD cohort and close friends for your laughter and support along the way: Vicky Morgan, Kelsey Lyberger, Anita To, Yige Luo, Katherine Corn, Mikaela Provost, Caitlin French, Charlotte Pickett, Matt van Avermaete, Hayley Rousek and Debbie Pattison.

Chapter 1 would not have been completed without the generosity of my collaborators: Graham Coop and Santiago Ramirez for believing in my vision for a cross-continent comparative study and welcoming me into their labs; and Marcelo Nicolás Agra and María Alejandra Palacio for helping me realize this vision with honey bee samples from Argentina. Thank you to Daniel Gates, Anne Lorant, M. Taylor Perkins, Graham Coop and Jeff Ross-Ibarra for their advice and help with Chapter 2.

This thesis would not have been completed without the love and support of my family. Thank you to my parents, Kat and Dave, and sister Alexa for all of your encouragement, and my cat Houdini for being at my side the whole way. Lastly, thank you to my husband Joshua Foster for reminding me there are many mountains to climb (including my PhD) and I can summit any of them if I just keep going up.