\documentclass[12pt]{ucdavisthesis}

% PLEASE READ THE MANUAL - ucdavisthesis.pdf (in the package installation directory)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
%               LATEX COMMANDS FOR DOCUMENT SETUP                      %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\usepackage{bookmark}
\usepackage[us,nodayofweek,12hr]{datetime}
\usepackage{graphicx}

\RequirePackage[font=small,labelfont=bf]{caption}
\RequirePackage{amsmath,amssymb,amsthm}
\RequirePackage{graphicx}
\RequirePackage[hidelinks]{hyperref}
\RequirePackage{subcaption}
\RequirePackage{tabularx}
\RequirePackage{wasysym}
\RequirePackage{authblk}
\RequirePackage{bm}
\RequirePackage{bbm}
\RequirePackage{color}
\RequirePackage{nicefrac}

\RequirePackage{lineno}
\RequirePackage{mathtools}
\RequirePackage[bibstyle=authoryear,citestyle=authoryear-comp,
                date=year,
                maxbibnames=9,maxnames=5,maxcitenames=2,
                backend=biber,uniquelist=false,uniquename=false,
                % style=apa,
                sorting=nyt,
                hyperref=true]{biblatex}

\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\V}{\text{V}}

\newcommand{\graham}[1]{\todo[size=\scriptsize, color=red!50]{#1}}
\newcommand{\vince}[1]{\todo[size=\scriptsize, color=blue!50]{#1}}

\newcommand{\cf}{\emph{cf.} }
\DeclareMathOperator{\var}{Var}
\DeclareMathOperator{\cov}{Cov}
\DeclareMathOperator{\T}{{\mathrm{T}}}
\newcommand{\vect}[1]{\mathbf{#1}}
\newcommand{\nssh}{SSH_n}
\newcommand{\nnn}{\nonumber}
\newcommand{\chapquote}[2]{\begin{quotation} \textit{#1} \end{quotation} \begin{flushright} - #2\end{flushright} }

\DeclareMathOperator{\flt}{\mathrm{flat}}

\newcommand{\msxa}{\mars_X}
\newcommand{\fsxa}{\venus_X}

\bibliography{biblio}

\hyphenation{dis-ser-ta-tion blue-print man-u-script pre-par-ing} %add hyphenation rules for words TeX doesn't know


%\renewcommand{\rightmark}{\scriptsize A University of California Davis\ldots \hfill Rev.~\#1.0 \quad Compiled: \currenttime, \today}
% a fancier running header that can be used with draftcls options

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
%        DOCUMENT SETUP AND INFORMATION FOR PRELIMINARY PAGES          %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title          {Exploring the Effects of Linkage for Ancestry and Temporal Polygenic Adaptation}
%Exact title of your thesis. Indicate italics where necessary by underlining or using italics. Please capitalize the first letter of each word that would normally be capitalized in a title.

\author         {Vincent Scott Buffalo}
%Your full name as it appears on University records. Do not use initials.

\authordegrees  { \;\;}
%Indicate your previous degrees conferred.

\officialmajor  {Population Biology}
%This is your official major as it appears on your University records.

\graduateprogram{Population Biology Graduate Group}
%This is your official graduate program name. Used for UMI abstract.

\degreeyear     {2019}
% Indicate the year in which your degree will be officially conferred.

\degreemonth    {September}
% Indicate the month in which your degree will be officially conferred. Used for UMI abstract.

\committee{Graham M. Coop}{Charles H. Langley}{Michael Turelli}{David J. Begun}{}
% These are your committee members. The command accepts up to five committee members so be sure to have five sets of braces, even if there are empties.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\copyrightyear{2019}
\nocopyright

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\dedication{\textsl{To ____ .}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\abstract{

Gene flow between diverged populations is often deleterious (e.g. due to genetic incompatibilities or divergent ecologies),
but introgression can also be a source of rapid evolution, especially to novel niches.
My thesis focuses on the evolutionary outcomes of these opposing selection forces on introgressed ancestry,
the repeatability of these outcomes across admixed populations,
and the consequences for species niches and adaptive limitations.
I have developed novel population genetics methods to detect selection in admixed populations,
and applied these methods to disentangle how demography and selection have shaped the evolution
and range expansions of two very different species: textit{Scutellata}-European hybrid honey bees and highland maize.

For my first dissertation chapter, I conducted a cross-continental comparison of the outcomes of admixture and selection in \textit{Scutellata}-European honey bees.
\textit{Scutellata} honey bees from South Africa were introduced to the Americas in the 1950s, but soon escaped, and through interbreeding with European honey bees,
formed a highly successful and invasive hybrid population that spread at ~300km/year across the Americas.
This is a great system to study the adaptive potential and limitations of admixture in facilitating rapid range expansions,
with natural replication in North and South America.
For this research, I collected and sequenced 300+ bees from two nearly 1000 km transects,
one in California and one in Argentina, to compare ancestry patterns across their genomes.
I found evidence of convergent selection favoring African \textit{Scutellata} honey bee ancestry at a number of loci in the genome in North and South America.
These loci are strong candidates for contributing to the high fitness and success of \textit{Scutellata}-European hybrid honey bees.
Because these bees are highly defensive, their continued spread is an agricultural and public health concern.
I found parallel clines in genomewide ancestry between continents at similar latitudes,
despite much larger dispersal distances to reach California from the origin of the invasion,
evidence that many loci across the genome are currently preventing spread to temperate zones.

For my second dissertation chapter, I analyzed the outcome of introgression between maize and its wild highland 'teosinte' relative, \textit{mexicana},
which may have facilitated maize’s range expansion from the valleys where it was domesticated up to 3000m in the mountains of Mexico.
Using a novel method that accounts for background patterns of ancestry variance and covariance between populations
(e.g. due to gene flow or shared drift post-admixture), I found strong evidence for adaptive introgression from \textit{mexicana} into maize,
especially among the highest elevation populations, consistent with introgression facilitating maize’s colonization of the highlands.
I also found loci (including a newly identified inversion) where selection maintains steep ancestry clines across elevation.
I demonstrated evidence of selection against introgression,
removing \textit{mexicana} ancestry from near domestication genes and lower recombination regions of the genome (due to linked selection).
One surprising finding is that despite observations of hybrids in the field,
and opportunities for gene flow from locally adapted \textit{mexicana} that grows side-by-side with contemporary maize,
I found little evidence for recent locally sourced haplotypes genomewide or at loci with high local introgression.
Rather, the majority of introgression is from >1000 generations ago,
and has subsequently diverged within the maize background and been sorted by selection along an elevational cline and within individual populations (see Calfee et al. 2021 bioRxiv).
This work has broader impacts for understanding the longer term effects of introgression on range expansions and aiding in the discovery of key loci associated with high-elevation adaptations,
which may be crucial for future breeding of maize, a global staple, under climate change.

Overall, this thesis adds to our knowledge of the adaptive potential (and limitations) of introgression in range and niche expansions,
and provides in-depth genomic analyses of selection and admixture in two agriculturally-relevant species.

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\acknowledgments{
  Thank you everyone.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Each chapter can be in its own file for easier editing and brought in with the \include command.
% Then use the \includeonly command to speed compilation when working on a particular chapter.
%%% \includeonly{ucdavisthesis_example_Chap1}

\begin{document}

%\newcommand{\bibfont}{\singlespacing}
% need this command to keep single spacing in the bibliography when using natbib

%\bibliographystyle{unsrtnat}
%many other bibliography styles are available (IEEEtran, mla, etc.). Use one appropriate for your field.

\makeintropages %Processes/produces the preliminary pages

\include{chapter_1}
\include{chapter_2}


\begingroup
\setstretch{0.8}
\setlength\bibitemsep{0pt}

% citations from chapter 1
%\nocite{dplyr}
%\nocite{ggplot2}
%\nocite{R}



\printbibliography
\endgroup

\end{document}
